---
title: "Data Transformation and Spatial Distribution of SOC"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

### Normal Score Transformation:

Prior to statistical  and geostatistical analysis, the data of SOC were tested for normality by Shapiro-Wilk normality test.  We used normal score transformation to full fill multi-Guassian assumption necessary for geostatistical analysis. 


## Load packages

```{r}
library(plyr)
library(dplyr)
library(raster)
library(agricolae)
library(plyr)
library(ggplot2)
library(maptools)
library(gstat)
library(RColorBrewer)
library(classInt)
library(latticeExtra)
library(sp)
```

### Set working directory

```{r}
setwd("D:\\SOC_paper\\Data")
dir ()
```

### Load data

```{r}
mf<-read.csv("GOPAL_SOIL.csv",header=TRUE)          # All data
bound<-readShapePoly("GOPAL_BD.shp")                # District boundary 
names(mf)
```

### Data transformation

```{r}
source("nscore.R")
SOC.ns <- nscore(mf$SOC)
mf$SOC.ns<- SOC.ns$nscore
```

### Shapiro-Wilk test of normality

```{r}
shapiro.test(mf$SOC)
shapiro.test(mf$SOC.ns)
```

### Histogram SOC and SOC.ns

```{r}
hist (mf$SOC)
hist (mf$SOC.ns)
```

### Import  As TIF file

#### Histogram: Soil OC (SOC)

```{r}
windows(width=3.25, height=3.25)
tiff(file="FIG_SOC_HISTO.tif", 
     width=3.25, 
     height=3.25, 
     units = "in", 
     pointsize = 12, 
     res=600, 
     bg = "white",
     restoreConsole = T,
     compression =  "lzw")
hist(mf$SOC, main="", 
     xlab= "", 
     ylab="",
     cex.main=2.5)
dev.off()
```

#### Histogram: Normal Score Transformed (SOC.ns)

```{r}
windows(width=3.25, height=3.25)
tiff(file="FIG_SOC_ns_HISTO.tif", 
     width=3.25, 
     height=3.25, 
     units = "in", 
     pointsize = 12, 
     res=600, 
     bg = "white",
     restoreConsole = T,
     compression =  "lzw")
hist(mf$SOC.ns, main="", 
     xlab= " ", 
     ylab="",
     cex.main=2.5)
dev.off()
```

## Create Calibration and Validation Data Set

```{r}
# proportion of full dataset to use for training (70%)
tr_prop = 0.70
# training data
mfC = ddply(mf, .(AEZ,GEO,LT,SP),
    function(., seed) { set.seed(seed); .[sample(1:nrow(.), trunc(nrow(.) * tr_prop)), ] }, seed = 101)   #calibration dataset
# Validation data (30% of data)
mfV = ddply(mf,  .(AEZ,GEO,LT,SP), 
    function(., seed) { set.seed(seed); .[-sample(1:nrow(.), trunc(nrow(.) * tr_prop)), ] }, seed = 101)  #validation dataset
write.csv(mfV, "GOPAL_Validation.csv")
write.csv(mfC, "GOPAL_Calibration.csv")
```

### Desciptive Statistics

#### Desciptive Statistics: All data

```{r}
length(mf$SOC)                # Number of observation
summary(mf$SOC)               # Summary stat
sd(mf$SOC)                    # Standard deviation
skewness(mf$SOC)              # skewness
sd(mf$SOC)*100/mean(mf$SOC)   # CV 
```


#### Desciptive Statistics: Calibration data

```{r}
length(mfC$SOC)                  # Number of observation
summary(mfC$SOC)                 # Summary stat
sd(mfC$SOC)                      # Standard deviation
skewness(mfC$SOC)                # Skewness
sd(mfC$SOC)*100/mean(mfC$SOC)    # CV 
```

#### Desciptive Statistics: Validation data

```{r}
length(mfV$SOC)                 # Number of observation
summary(mfV$SOC)                # Summary stat
sd(mfV$SOC)                     # Standard deviation
skewness(mfV$SOC)               # Skewness
sd(mfV$SOC)*100/mean(mfV$SOC)   # CV 
```

### Maps Observed SOC with Calibartion and Validation data 

```{r}
coordinates(mfV)<-~x+y
at.soc = classIntervals(mf$SOC, n = 10, style = "quantile")$brks
round(quantile(mf$SOC, probs=seq(0,1, by=0.1)),1)
rgb.palette.col <- colorRampPalette(c("red","yellow", "green", "blue"),space = "rgb")

soc<-levelplot(SOC~x+y, mf,cex=0.6,
              aspect = "iso",main= "",
              xlab="", ylab="",
              xlim=c(773500, 825400),ylim=c(2525000,2586000 ),
              scales=list(y=list(draw=T,cex=0.5,rot=90, tck= 0.5),x=list(draw=T, cex=0.5,tck= 0.6)),
              par.settings=list(axis.line=list(col="grey",lwd=0.5)),
              col.regions=rgb.palette.col (20),at=at.soc,
              colorkey=list(space="right",width=1.2,at=1:11,labels=list(cex=1.2,at=1:11,
              labels=c("" ,"<1.2", "", "1.8" ,"", "2.6","", "4.2", "" , ">7.8", ""))),
              panel = function(...) {
              panel.levelplot.points(...)
              sp.points(mfV, col="black", cex=1.0,pch=21)
              sp.polygons(bound,lty=1,lwd=0.5,col="grey30")
              },) 

windows(width=5, height=5)
tiff(file="FIG_SOC_OBSERVED.tif", 
width=5, height=5,units = "in", pointsize = 12, res=1600,
restoreConsole = T,bg="transparent")
print(soc)
dev.off()
soc
```


## Distribution of SOC in  AEZs

#### AEZ mean

```{r}
aez.stat<-ddply(mf, "AEZ", summarise,
      N=length(SOC), Mean = mean(SOC), SD = sd(SOC), Min=min(SOC),Max=max(SOC))
aez.stat
write.csv(aez.stat, "AEZ_STAT.csv")
```

#### AEZ:  ANOVA 

```{r}
aez<-aov(SOC~AEZ, mf)
summary(aez)
```

#### AEZ: LSD test

```{r}
LSD.test(aez, "AEZ",console=TRUE)
```

#### AEZ: Boxplot

```{r}
windows(width=7.5, height=5)
tiff( file="FIG_BOX_PLOT_SOC_AEZ.tif", 
width=7.5, height=5,units = "in", pointsize = 12, res=300, bg = "transparent",
restoreConsole = TRUE)
rgb.palette <- colorRampPalette(c("red","yellow","green", "blue", "skyblue"),
space = "rgb")

p=ggplot(mf, aes(y=SOC, x=AEZ)) +
         geom_point(aes(colour=SOC),size = I(2.0),
         position=position_jitter(width=0.05, height=0.05)) +
         geom_boxplot(fill=NA, outlier.colour=NA) +
         labs(title="")+
         theme_bw() +
         coord_flip()+
            theme(axis.line = element_line(colour = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.text.y=element_text(size=10),
            axis.text.x = element_text(size=12))
p1= p+scale_colour_gradientn(name="Soil OC \n(%)", colours =rgb.palette(20),na.value = "transparent")
p2=p1 + theme(legend.text = element_text(size = 10),legend.title = element_text(size = 12))
p2+ labs(y="Soil OC(%)", x = "")
dev.off()
p2+ labs(y="Soil OC(%)", x = "")
```


## Distribution of SOC by  Surface geology

#### GEO: descriptive stat

```{r}
geo.stat<-ddply(mf, "GEO", summarise,
      N=length(SOC), Mean = mean(SOC), SD = sd(SOC), Min=min(SOC),Max=max(SOC))
geo.stat
write.csv(geo.stat, "GEO_STAT.csv")
```

#### GEO: ANOVA 

```{r}
geo <- aov(SOC~GEO, mf)
summary(geo)
```

#### GEO: LSD test

```{r}
LSD.test(geo, "GEO",console=TRUE)
```

#### GEO: Boxplot

```{r}
windows(width=7.5, height=5)
tiff( file="FIG_BOX_PLOT_SOC_GEO.tif", 
width=7.5, height=5,units = "in", pointsize = 12, res=300, bg = "transparent",
restoreConsole = TRUE)
rgb.palette <- colorRampPalette(c("red","yellow","green", "blue", "skyblue"),
space = "rgb")

p=ggplot(mf, aes(y=SOC, x=GEO)) +
         geom_point(aes(colour=SOC),size = I(2.0),
         position=position_jitter(width=0.05, height=0.05)) +
         geom_boxplot(fill=NA, outlier.colour=NA) +
         labs(title="")+
         theme_bw() +
         coord_flip()+
            theme(axis.line = element_line(colour = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.text.y=element_text(size=10),
            axis.text.x = element_text(size=12))
p1= p+scale_colour_gradientn(name="Soil OC \n(%)", colours =rgb.palette(20),na.value = "transparent")
p2=p1 + theme(legend.text = element_text(size = 10),legend.title = element_text(size = 12))
p2+ labs(y="Soil OC(%)", x = "")
dev.off()
p2+ labs(y="Soil OC(%)", x = "")
```


## Distribution of SOC by  landtype

#### Landtype: descriptive stat

```{r}
lt.stat<-ddply(mf, "LT", summarise,
      N=length(SOC), Mean = mean(SOC), SD = sd(SOC), Min=min(SOC),Max=max(SOC))
lt.stat
write.csv(lt.stat,"LT_STAT.csv")
```

#### Landtype: ANOVA 

```{r}
lt<-aov(SOC~LT, mf)
summary(lt)
```

#### Landtype: LSD test

```{r}
LSD.test(lt, "LT",console=TRUE)
```

#### Landtype: Boxplot

```{r}
windows(width=7.5, height=5)
tiff( file="FIG_BOX_PLOT_SOC_LT.tif", 
width=7.5, height=5,units = "in", pointsize = 12, res=300, bg = "transparent",
restoreConsole = TRUE)
rgb.palette <- colorRampPalette(c("red","yellow","green", "blue", "skyblue"),
space = "rgb")

p=ggplot(mf, aes(y=SOC, x=LT)) +
         geom_point(aes(colour=SOC),size = I(2.0),
         position=position_jitter(width=0.05, height=0.05)) +
         geom_boxplot(fill=NA, outlier.colour=NA) +
         labs(title="")+
         theme_bw() +
         coord_flip()+
            theme(axis.line = element_line(colour = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.text.y=element_text(size=10),
            axis.text.x = element_text(size=12))
p1= p+scale_colour_gradientn(name="Soil OC \n(%)", colours =rgb.palette(20),na.value = "transparent")
p2=p1 + theme(legend.text = element_text(size = 10),legend.title = element_text(size = 12))
p2+ labs(y="Soil OC(%)", x = "")
dev.off()
p2+ labs(y="Soil OC(%)", x = "")
```


## Distribution of SOC in drainage class

#### Driange: descriptive stat

```{r}
dc.stat<-ddply(mf, "DC", summarise,
      N=length(SOC), Mean = mean(SOC), SD = sd(SOC), Min=min(SOC),Max=max(SOC))
dc.stat
write.csv(dc.stat,"DC_STAT.csv")
```

#### Drainage: ANOVA 

```{r}
dc<-aov(SOC~DC, mf)
summary(dc)
```

#### Drainage: LSD test

```{r}
LSD.test(dc, "DC",console=TRUE)
```

#### Drainage: Boxplot

```{r}
windows(width=7.5, height=5)
tiff( file="FIG_BOX_PLOT_SOC_DC.tif", 
width=7.5, height=5,units = "in", pointsize = 12, res=300, bg = "transparent",
restoreConsole = TRUE)
rgb.palette <- colorRampPalette(c("red","yellow","green", "blue", "skyblue"),
space = "rgb")

p=ggplot(mf, aes(y=SOC, x=DC)) +
         geom_point(aes(colour=SOC),size = I(2.0),
         position=position_jitter(width=0.05, height=0.05)) +
         geom_boxplot(fill=NA, outlier.colour=NA) +
         labs(title="")+
         theme_bw() +
         coord_flip()+
            theme(axis.line = element_line(colour = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.text.y=element_text(size=10),
            axis.text.x = element_text(size=12))
p1= p+scale_colour_gradientn(name="Soil OC \n(%)", colours =rgb.palette(20),na.value = "transparent")
p2=p1 + theme(legend.text = element_text(size = 10),legend.title = element_text(size = 12))
p2+ labs(y="Soil OC(%)", x = "")
dev.off()
p2+ labs(y="Soil OC(%)", x = "")
```



## Distribution of SOC in Soil permeability

#### Soil permeability: descriptive stat

```{r}
sp.stat<-ddply(mf, "SP", summarise,
      N=length(SOC), Mean = mean(SOC), SD = sd(SOC), Min=min(SOC),Max=max(SOC))
sp.stat
write.csv (sp.stat,"SP_STAT.csv")
```

#### Soil permeability: ANOVA 

```{r}
sp<-aov(SOC~SP, mf)
summary(sp)
```

#### Soil permeability: LSD test

```{r}
LSD.test(sp, "SP",console=TRUE)
```

#### Soil permeability: Boxplot

```{r}
windows(width=7.5, height=5)
tiff( file="FIG_BOX_PLOT_SOC_SP.tif", 
width=7.5, height=5,units = "in", pointsize = 12, res=300, bg = "transparent",
restoreConsole = TRUE)
rgb.palette <- colorRampPalette(c("red","yellow","green", "blue", "skyblue"),
space = "rgb")

p=ggplot(mf, aes(y=SOC, x=SP)) +
         geom_point(aes(colour=SOC),size = I(2.0),
         position=position_jitter(width=0.05, height=0.05)) +
         geom_boxplot(fill=NA, outlier.colour=NA) +
         labs(title="")+
         theme_bw() +
         coord_flip()+
            theme(axis.line = element_line(colour = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.text.y=element_text(size=10),
            axis.text.x = element_text(size=12))
p1= p+scale_colour_gradientn(name="Soil OC \n(%)", colours =rgb.palette(20),na.value = "transparent")
p2=p1 + theme(legend.text = element_text(size = 10),legend.title = element_text(size = 12))
p2+ labs(y="Soil OC(%)", x = "")
dev.off()
p2+ labs(y="Soil OC(%)", x = "")
```

## Distribution of SOC in diffrent landuse

#### Landuse: descriptive stat

```{r}
lu.stat<-ddply(mf, "LU", summarise,
      N=length(SOC), Mean = mean(SOC), SD = sd(SOC), Min=min(SOC),Max=max(SOC))
lu.stat
write.csv(lu.stat,"LU_STAT.csv")
```

#### Landuse: ANOVA 

```{r}
lu <- aov(SOC~LU, mf)
summary(lu)
```

#### Landuse: LSD test

```{r}
LSD.test(lu, "LU",console=TRUE)
```

#### Soil permeability: Boxplot

```{r}
windows(width=7.5, height=5)
tiff( file="FIG_BOX_PLOT_SOC_LU.tif", 
width=7.5, height=5,units = "in", pointsize = 12, res=300, bg = "transparent",
restoreConsole = TRUE)
rgb.palette <- colorRampPalette(c("red","yellow","green", "blue", "skyblue"),
space = "rgb")

p=ggplot(mf, aes(y=SOC, x=LU)) +
         geom_point(aes(colour=SOC),size = I(2.0),
         position=position_jitter(width=0.05, height=0.05)) +
         geom_boxplot(fill=NA, outlier.colour=NA) +
         labs(title="")+
         theme_bw() +
         coord_flip()+
            theme(axis.line = element_line(colour = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.text.y=element_text(size=10),
            axis.text.x = element_text(size=12))
p1= p+scale_colour_gradientn(name="Soil OC \n(%)", colours =rgb.palette(20),na.value = "transparent")
p2=p1 + theme(legend.text = element_text(size = 10),legend.title = element_text(size = 12))
p2+ labs(y="Soil OC(%)", x = "")
dev.off()
p2+ labs(y="Soil OC(%)", x = "")
```

### Correlation between SOC and DEM

```{r}
summary (mf$DEM)
cor.test(mf$SOC, mf$DEM)
```

### Correlation between SOC and NDVI

```{r}
summary (mf$NDVI)
cor.test(mf$SOC, mf$NDVI)
```

