---
title: 'Digital Mapping of Soil Properties: Soil Organic Carbon'
output:
  word_document: default
  pdf_document: default
  html_document: default
---

### Required Packages

```{r message=F, warning=F}
library(caret)
library(caretEnsemble)
library(doParallel)
library(randomForest)
library(plyr)
library(gstat)
library(rgdal)
library(raster)
library(RColorBrewer)
library(classInt)
library(latticeExtra)
library(maptools)
library(sp)
library(car)
library(rfUtilities)
```

### Set a working dirctory

```{r}
setwd("D:\\SOC_paper\\Data")
```

#### Import data

```{r}
grid <- read.csv("GOPAL_GRID.csv",header=TRUE)          # grid data
mf <- read.csv("GOPAL_SOIL.csv",header=TRUE)            # Data
bound <- readShapePoly("GOPAL_BD.shp")                  # District boundary 
polys <- list("sp.lines", as(bound, "SpatialLines"), col="black", lwd=.8,lty=1)
```

#### Covnvert catagorical variables to factors

```{r}
# Soil data
as.factor(mf$GEO_ID)-> mf$GEO_ID
as.factor(mf$AEZ_ID)-> mf$AEZ_ID
as.factor(mf$LT_ID)-> mf$LT_ID
as.factor(mf$DC_ID)-> mf$DC_ID
as.factor(mf$SP_ID)-> mf$SP_ID
as.factor(mf$LU_ID)-> mf$LU_ID

# Prediction grid
as.factor(grid$GEO_ID)-> grid$GEO_ID
as.factor(grid$AEZ_ID)-> grid$AEZ_ID
as.factor(grid$LT_ID)-> grid$LT_ID
as.factor(grid$DC_ID)-> grid$DC_ID
as.factor(grid$SP_ID)-> grid$SP_ID
as.factor(grid$LU_ID)-> grid$LU_ID
```

# Normal Score Transformation

```{r}
source("nscore.R")
SOC.ns <- nscore(mf$SOC)
mf$SOC.ns<- SOC.ns$nscore
```

## Create Calibration and Validation Data Set

```{r}
# proportion of full dataset to use for training (70%)
tr_prop = 0.70
# training data
mfC = ddply(mf, .(AEZ_ID,LT_ID,LU_ID),function(., seed) { set.seed(seed); .[sample(1:nrow(.), trunc(nrow(.) * tr_prop)), ] }, seed = 101)
# Validation data (30% of data)
mfV = ddply(mf,  .(AEZ_ID,LT_ID,LU_ID), 
    function(., seed) { set.seed(seed); .[-sample(1:nrow(.), trunc(nrow(.) * tr_prop)), ] }, seed = 101)
write.csv(mfV, "GOPAL_Validation.csv")
write.csv(mfC, "GOPAL_Calibration.csv")
length(mfC$x)
length(mfV$x)
```

#### Response variables: SOC

```{r}
V1 <- mfC$SOC.ns 
```

#### Random Forest Regression Model

```{r}
set.seed(1258)
fit.rf = randomForest(x = mfC[c(14:21)], y = V1, 
       importance=TRUE,  mtry=2, ntree=5001,maxnodes=3,type=regression) 
```

#### RF regression results 

```{r}
print(fit.rf)
fit.rf.results <- fit.rf$results
write.csv(fit.rf.results, "RF_SOC_STAT.csv")
```

#### Variables Importance

```{r}
varImpPlot(fit.rf,type=1, main="", cex=1, pch=19)
imp.mean=importance(fit.rf, type=1)
imp.sd=fit.rf$importanceSD
write.csv(imp.mean, "MEAN_IMPORTANCE_SOC.csv")
write.csv(imp.sd, "SD_IMPORTANCE_SOC.csv")
```

### RF prediction at validation data points

```{r}
mfV$RF_TREND <- predict(fit.rf, mfV[c(14:21)])
```
```{r}
names(mfV)
```
#### RF validation Statistics

```{r}
# RF root mean square error (RMSE)
RMSE_RF=sqrt(mean( (mfV$SOC.ns-mfV$RF_TREND)^2 , na.rm = TRUE ))
RMSE_RF
# RF R-squared
summary(lm(SOC.ns~RF_TREND, mfV))$r.squared
summary(lm(SOC.ns~RF_TREND, mfV))
```
# mean absolute error (MAE)
```{r}
MAE_RF=mean(abs(mfV$SOC.ns-mfV$RF_TREND))
MAE_RF
```
# write cross validation stat
```{r}
error<-rbind(RMSE_RF,MAE_RF)
write.csv(error, "Cross_Validation_STAT_RF_SOC.ns.csv")
```

### RF regression at prediction grid (Trend)

```{r}
grid$RF_TREND <-  predict(fit.rf, grid[c(4:11)])
```

#### Creat RF regression trend raster (geoTIF)

```{r,warning=FALSE}
rf.r <- rasterFromXYZ(as.data.frame(grid)[, c("x", "y", "RF_TREND")])
plot(rf.r, main="RF SOC Trend Map")
writeRaster(rf.r,filename="RF_SOC_TREND.tiff","GTiff",overwrite=TRUE)
```

## Random Forest Kriging (RFK)

#### Calcualte RF Residuals

```{r}
mfC$RF_TREND <- predict(fit.rf)                   # Predict 
mfC$RF_RES <- (mfC$SOC.ns-mfC$RF_TREND)           # Residuals
```

#### Spatial Prediction RF residuals

Before spatial prediction, we have to convert GPc, GPv and GPg data to spatial point data frame

```{r}
coordinates(mfC)<-~x+y
coordinates(mfV)<-~x+y
coordinates(grid)<-~x+y
```

#### Variogram of RF-residuals

```{r, warning=FALSE, fig.width = 4.5, fig.height = 4.5}
v.rf<-variogram(RF_RES ~ 1, data = mfC, cressie = T, cutoff=20000, width=20000/20)
plot(v.rf, pch=20, cex=1, col="Black")
m.v.rf<-vgm(1,"Exp",5000,3)
(m.v.f.rf<-fit.variogram(v.rf, m.v.rf))
plot(v.rf, pl=F, model=m.v.f.rf,col="black", cex=1, lwd=0.5,lty=1,pch=20,
main="Variogram of RF-Residuals ",xlab="Distance (m)",ylab="Semivariance")
write.csv(v.rf, "SOC_RES_vario.csv")
```

#### Simple Kriging (SK) Validation of RF-residuals

```{r}
sk.v.rf<-krige(RF_RES ~ 1, mfC,mfV, model=m.v.f.rf, beta=0) 
mfV$RF_SK<-sk.v.rf$var1.pred
```

#### Calculate RFK values for validation data set 

```{r}
mfV$RFK<-mfV$RF_SK+mfV$RF_TREND
write.csv(mfV, "RF_validation_results.csv")
```

#### Validation statistics of RFK

```{r}
# RFK root mean square error (RMSE)
RMSE_RFK=sqrt(mean( (mfV$SOC.ns-mfV$RFK)^2 , na.rm = TRUE ))
RMSE_RFK
# RFK R-squared
summary(lm (SOC.ns ~ RFK, mfV)) $r.squared
summary(lm (SOC.ns ~ RFK, mfV))

```
# Mean absolute error (MAE)
```{r}
MAE_RFK=mean(abs(mfV$SOC.ns-mfV$RFK))
MAE_RFK
```
# Cross validation stat
```{r}
error<-rbind(RMSE_RFK,MAE_RFK)
write.csv(error, "Cross_Validation_STAT_RFK_SOC.ns.csv")
```

#### Simple kirging of RF residuals at prediction girds

```{r}
sk.rf<-krige(RF_RES~ 1, mfC,grid, model=m.v.f.rf, beta=0)
grid$RF_SK<-sk.rf$var1.pred
sk.rf.r <- rasterFromXYZ(as.data.frame(grid)[, c("x", "y", "RF_SK")])
writeRaster(sk.rf.r,filename="RF_SOC_SK_Residuals.tiff","GTiff",overwrite=TRUE)
```

#### Map of RF-residuals

```{r fig.width = 6, fig.height = 6}
plot(sk.rf.r, main="RF SOC Residuals Map")
```

#### Final RFK kriging Prediction Map (RF-Trend+RF-residuals)

```{r,warning=FALSE}
grid$RFK<-grid$RF_TREND+grid$RF_SK
```

# Back trasformation

```{r}
grid$RFK.bt <- backtr(grid$RFK,SOC.ns, tails='separate') # Back transformation
x<-quantile(mf$SOC,probs= 0.99,na.rm=TRUE)
grid[grid$RFK.bt >x] <- x  
rfk.r <- rasterFromXYZ(as.data.frame(grid)[, c("x", "y", "RFK.bt")])
```

#### Plot RFK map

```{r fig.width = 6, fig.height = 6}
plot(rfk.r, main="RFK SOC Prediction Map ")
```

#### Write RFK raster map

```{r}
writeRaster(rfk.r,filename="RFK_SOC_Prediction.tiff","GTiff",overwrite=TRUE)
```

## Ordinary kriging (OK)

#### Varigraom of SOC

```{r, warning=FALSE, fig.width = 4.5, fig.height = 4.5}
v<-variogram(SOC.ns ~ 1, data = mfC, cutoff=20000, width=20000/20,cressie = T)
plot(v, pch=20, cex=1, col="Black")
m.v<-vgm(1,"Exp",5000,3)
(m.v.f<-fit.variogram(v, m.v))
plot(v, pl=F, model=m.v.f,col="black", cex=1, lwd=0.5,lty=1,pch=20,
main="Variogram of SOC ",xlab="Distance (m)",ylab="Semivariance")
write.csv(v, "SOC_vario.csv")
```


#### OK Validation

```{r}
ok.v<-krige(SOC.ns ~ 1, mfC,mfV, model=m.v.f) 
mfV$OK<-ok.v$var1.pred
write.csv(mfV, "OK_Validation.csv")
```

#### Validation statistics of oK

```{r}
# OK root mean square error (RMSE)
RMSE_OK=sqrt(mean( (mfV$SOC.ns-mfV$OK)^2 , na.rm = TRUE ))
RMSE_OK
# OK R-squared
summary(lm(SOC.ns~OK, mfV))$r.squared
summary(lm(SOC.ns~OK, mfV))
```
# Mean absolute error
```{r}
MAE_oK=mean(abs(mfV$SOC.ns-mfV$OK))
MAE_oK
```
# Cross validation stat
```{r}
error<-rbind(RMSE_OK,MAE_oK)
write.csv(error, "Cross_Validation_STAT_OK_SOC.ns.csv")
```

#### OK prediction

```{r,warning=FALSE}
ok<-krige(SOC.ns ~ 1, mfC,grid, model=m.v.f)
grid$OK<-ok$var1.pred
ok.r <- rasterFromXYZ(as.data.frame(grid)[, c("x", "y", "OK")])
```

#### Back-transformation

```{r}
grid$OK.bt <- backtr(grid$OK,SOC.ns, tails='separate') # Back transormation
x<-quantile(mf$SOC,probs= 0.99,na.rm=TRUE)
grid[grid$OK.bt >x] <- x 
ok.r <- rasterFromXYZ(as.data.frame(grid)[, c("x", "y", "OK.bt")])
```

#### OK Plot OK map

```{r fig.width = 6, fig.height = 6}
plot(ok.r, main="OK Prediction SOC (%) ")
```

#### Write OK raster map

```{r,warning=FALSE}
writeRaster(ok.r,filename="OK_SOC_Prediction.tiff","GTiff",overwrite=TRUE)
```

#### Write Final Prediction RFK-OK output

```{r}
write.csv(grid, "RFK_OK_Prediction.csv")
```

### Plot RFK and OK Map

```{r fig.width = 8, fig.height = 5}
gridded(grid) <- TRUE
polys<- list("sp.lines", as(bound, "SpatialLines"), col="black", lwd=.8,lty=1)
rgb.palette <- colorRampPalette(c("blue", "green",  "yellow", "orange","red"),space = "rgb")
spplot(grid, c("OK.bt", "RFK.bt"), 
	names.attr = c("Ordinary Kriging (OK)", "Random Forest Kriging (RFK)"), 
	main = "", as.table = TRUE,
	sp.layout=list(polys),col.regions=rgb.palette(100) )
```


#### Plot SOC and Residuals Variograms

```{r, warning=FALSE, fig.width = 5.5, fig.height = 5.5}
v.ok<-read.csv("SOC_vario.csv",header=TRUE)
class(v.ok) <- "gstatVariogram"
v.r<-read.csv("SOC_RES_vario.csv",header=TRUE)
class(v.r) <- "gstatVariogram"

plot(v.ok$gamma ~ v.ok$dist, xlim = c(0, max(v.ok$dist) * 1.05),
ylim = c(0,1.2), pch = 20, col = "black", main = "",
cex.axis=1.1, xlab = "Distance (m)", ylab = "Semivariance")
points(v.r$gamma ~ v.r$dist,pch = 21, col = "black", cex = .55)
lines(variogramLine(fit.variogram(v, vgm(0.359,"Exp", 8425.03, 0.6481)), 
   maxdist = max(v$dist)), col = "black")
lines(variogramLine(fit.variogram(v.rf, vgm(0.4098,"Exp", 3933.97, 0.1707)), 
   maxdist = max(v.rf$dist)), col = "black")
```
